---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# SpecPP

<!-- badges: start -->
<!-- badges: end -->

A R package to conduct spectral analysis for multivariate spatial point pattern on rectangle region ([Ding et al., 2025](https://arxiv.org/abs/2502.09948)). Given a multivariate point pattern and a parametric model of the intensity function, this package enables you to depict the frequency domain feature of your data by kernel smoothing. Functions for bandwidth selection, visualization, and coherence analysis are provided in this package.

## Installation
You can install the development version of SpecPP from [GitHub](https://github.com/) with:

``` {r, eval=FALSE, echo = TRUE}
# install.packages("pak")
pak::pak("qwding101/SpecPP")
```

## Tutorial
TBA

## References
-   Ding, Q. W., Yang, J., & Shin, J. (2025). Pseudo-spectra of multivariate inhomogeneous spatial point processes. arXiv preprint [arXiv:2502.09948](https://arxiv.org/abs/2502.09948).
-   Yang, J., & Guan, Y. (2024). Fourier analysis of spatial point processes. arXiv preprint [arXiv:2401.06403](https://arxiv.org/abs/2401.06403).

<!-- You'll still need to render `README.Rmd` regularly, to keep `README.md` up-to-date. `devtools::build_readme()` is handy for this. In that case, don't forget to commit and push the resulting figure files, so they display on GitHub and CRAN. -->
