---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# fspp

<!-- badges: start -->
<!-- badges: end -->

A R package to conduct spectral analysis for multivariate spatial point pattern on rectangle region. Given a multivariate point pattern and parametric model of the intensity function, this package enables you to depict the frequency-domain feature of your data by kernel smoothing. Functions for bandwidth selection, visualization, and coherence analysis are provided in this package.

You can install the development version of fspp from [GitHub](https://github.com/) with:

``` {r, eval=FALSE, echo = TRUE}
# install.packages("pak")
pak::pak("qwding101/fspp")
```

-   For technical details, please read [our manuscript](https://arxiv.org/abs/2502.09948).
-   Package tutorial (under construction).
<!-- You'll still need to render `README.Rmd` regularly, to keep `README.md` up-to-date. `devtools::build_readme()` is handy for this. In that case, don't forget to commit and push the resulting figure files, so they display on GitHub and CRAN. -->
