---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# SpecPP

<!-- badges: start -->
<!-- badges: end -->

An R package for spectral analysis of multivariate spatial point patterns in a rectangular region ([Ding et al., 2025](https://arxiv.org/abs/2502.09948)). Given a multivariate point pattern and a parametric model of the intensity function, this package facilitates the characterization of frequency-domain features from your data through kernel smoothing. It includes functions for spectral density estimation, bandwidth selection, visualization, and coherence analysis.

## Installation
You can install the development version of SpecPP from [GitHub](https://github.com/) with:

``` {r, eval=FALSE, echo = TRUE}
# install.packages("pak")
pak::pak("qwding101/SpecPP")
```

## Tutorial
[This tutorial](https://qwding101.github.io/SpecPP/articles/lansing.html) demonstrates how to use SpecPP to analyze multivariate point pattern data.

## References
-   Ding, Q. W., Yang, J., & Shin, J. (2025). Pseudo-spectra of multivariate inhomogeneous spatial point processes. arXiv preprint [arXiv:2502.09948](https://arxiv.org/abs/2502.09948).
-   Yang, J., & Guan, Y. (2024). Fourier analysis of spatial point processes. arXiv preprint [arXiv:2401.06403](https://arxiv.org/abs/2401.06403).

<!-- You'll still need to render `README.Rmd` regularly, to keep `README.md` up-to-date. `devtools::build_readme()` is handy for this. In that case, don't forget to commit and push the resulting figure files, so they display on GitHub and CRAN. -->
