---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# SpecPP

<!-- badges: start -->

<!-- badges: end -->

This R package performs spectral analysis of multivariate spatial point patterns observed in a rectangular region ([Ding et al., 2025](https://arxiv.org/abs/2502.09948)). By specifying a parametric model for the intensity, it extracts the frequency‐domain characteristics of your data by smoothing the periodogram. Key features of this tool include spectral density estimation, cross-validated bandwidth selection, spectrum visualization, and coherence analysis.

## Installation

You can install the development version of SpecPP from [GitHub](https://github.com/) with:

```{r, eval=FALSE, echo = TRUE}
# install.packages("pak")
pak::pak("qwding101/SpecPP")
```

## FAQ

**Q:** What is a spatial point pattern?

**A:** Please refer to [link 1](https://geographicdata.science/book/notebooks/08_point_pattern_analysis.html) or [link 2](https://documentation.sas.com/doc/en/pgmsascdc/v_063/statug/statug_spp_overview02.htm) for an introduction to point patterns.

**Q:** How to use this R package?

**A:** [This tutorial](https://qwding101.github.io/SpecPP/articles/lansing.html) demonstrates how to use SpecPP to analyze multivariate point pattern data.

## References

-   Ding, Q. W., Yang, J., & Shin, J. (2025). Pseudo-spectra of multivariate inhomogeneous spatial point processes. arXiv preprint [arXiv:2502.09948](https://arxiv.org/abs/2502.09948).
-   Yang, J., & Guan, Y. (2024). Fourier analysis of spatial point processes. arXiv preprint [arXiv:2401.06403](https://arxiv.org/abs/2401.06403).

<!-- You'll still need to render `README.Rmd` regularly, to keep `README.md` up-to-date. `devtools::build_readme()` is handy for this. In that case, don't forget to commit and push the resulting figure files, so they display on GitHub and CRAN. -->
