---
title: "Analyzing Lansing Woods data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Analyzing Lansing Woods data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

In this tutorial, we use the Lansing Woods dataset to demonstrate how to extract frequency domain features from a multivariate inhomogeneous point pattern data using SpecPP package. Here, we focus on the usage of this package. For the interpretation of the result, please refer to Ding et al. (2025, Section 3.2, Section 8, and Appendix F.3).

# Import package and data

The Lansing Woods dataset, available in the `spatstat` package, contains the spatial distributions of six different tree species.

```{r setup, message=FALSE}
library(SpecPP)
library(spatstat)

plot(split(lansing), pch = 20, cex = .1, main = "Point pattern of Lansing Woods")
summary(lansing)
```

Before proceeding with the analysis, we first examine the side lengths of the observational window. If the window is too small, the spectral estimator will be evaluated on a coarse frequency grid, limiting its ability to capture the frequency-domain characteristics of the point pattern. By contrast, an excessively large window can make computations unfeasible. We recommend side lengths between 20 and 40, as they offer a good balance between computational efficiency and resolution.

As shown in `summary(lansing)`, the window is the unit square (since [`spatstat`](https://search.r-project.org/CRAN/refmans/spatstat.data/html/lansing.html) rescaled the original window). This is too small for spectral analysis, so we rescale it by a factor of 1/20:

```{r} 
scale.factor = 1/20
spp = rescale(lansing, scale.factor)
```

Now the rescaled dataset `spp` with side lengths 20 is suitable for our analysis.

```{r}
Window(spp)
```

# Estimate pseudo-spectrum

To estimate the pseudo-spectrum of `spp`, we use the kernel spectral density estimator (KSDE). This calculation of KSDE involves two key steps: intensity function estimation and bandwidth selection.

## Intensity function estimation

The KSDE involves the first-order intensity function, which needs to be estimated. While nonparametric approaches (e.g., `spatstat::density.ppp()`) are available, the asymptotic properties of the KSDE are established under the parametric form of the intensity. Therefore, for each tree species, we fit a log-linear model for the intensity function using Cartesian coordinates:

```{r}
spps = split(spp)
fit.lambda = vector("list", length(spps))
names(fit.lambda) = names(spps)
for (i in seq_along(spps)){
  fit.lambda[[i]] = predict.ppm(ppm(spps[[i]] ~ x + y + x:y))
}
```

We then plot the fitted intensity to verify whether it adequately captures spatial inhomogeneity. The figure below suggests that our model `~ x + y + x:y` may be overly simplistic. However, for pedagogical purposes, we proceed with this specification.

```{r}
plot(as.solist(fit.lambda), main = "Estimated intensity")
```

In practice, additional spatial covariates can be incorporated to improve the intensity model. These covariates should be specified in the model formula and provided through the `data` argument in `spatstat.model::ppm()`. For example, if you have two covariates `pH` and `gradient`, they should be combined into a list and passed to the `data` argument (check `?spatstat.model::ppm` for details on the format of the covariates):

```{r eval=FALSE}
# Just an example. This dataset from spatstat doesn't have any covariate.
covars = list(pH, gradient)
for (i in seq_along(spps)){
  fit.lambda[[i]] = predict.ppm(ppm(spps[[i]] ~ x + y + x:y + pH + gradient,
                                    data = covars))
}
```

## Bandwidth selection

Selecting a good bandwidth is crucial for reliable spectral estimation. This is handled by the `select_band()` function, which requires the following arguments:

-   `ppp`: The point pattern
-   `inten.formula`: The model for the intensity (mentioned in previous section), in our example, is `"~ x + y + x:y"`. Note that the formula here must be provided as a string.
-   `band.range`: A numeric vector defining the search space for the optimal bandwidth.

If additional spatial covariates are specified in `inten.formula`, the corresponding data must be provided via the `data.covariate` argument. Other arguments of `select_band()` can be left at their default values.

```{r}
cv = select_band(ppp = spp,
                 inten.formula = "~ x + y + x:y",
                 band.range = seq(0.7, 1.1, .01))

plot(x = cv$Result[1,], y = cv$Result[2,], type = "l",
     xlab = "Bandwidth", ylab = "Spectral divergence")
abline(v = cv$OptimalBandwidth, col = "blue", lwd = 2, lty = "dashed")
```

As you see in above figure, the `select_band()` calculates the spectral divergence across all candidate bandwidths in `band.range`. The bandwidth minimizing the spectral divergence is the optimal bandwidth we select, which is `cv$OptimalBandwidth` = `r cv$OptimalBandwidth` (marked by the blue dashed line).

### *Important considerations for bandwidth selection*

-   **Resolution of `band.range`**: The search space should be neither too fine nor too coarse to balance the computational efficiency with accuracy.

-   **Coverage of `band.range`**: A narrow search range may fail to capture the global minimum. When the minimum lies at the boundary of the specified range, `select_band()` issues a warning suggesting a broader search space. However, this safeguard is not foolproof—there are cases where no warning appears even if a local minimum is selected. For example, specifying `band.range = seq(0.96, 1.05, 0.01)` in this example fails to identify the global minimum, because the local minimum lies inside the search range.

To mitigate these risks, we recommend running `select_band()` multiple times with different `band.range`, visualizing the results to ensure the chosen bandwidth is robust and not constrained by an inadequate search space.

## Calculate the KSDE

Finally, we can calculate the KSDE with the fitted intensity and cross-validated bandwidth.

```{r}
KSDE = periodogram_smooth(ppp = spp,
                          inten.formula = "~ x + y + x:y",
                          bandwidth = cv$OptimalBandwidth)
```

The output of `periodogram_smooth()` is a list of `r length(spp)` (marginal) + `r length(spp)*(length(spp)-1)/2` (cross) = `r length(spp) + length(spp)*(length(spp)-1)/2` KSDE matrices.

```{r}
names(KSDE)
```

Note that the KSDE for cross pseudo-spectrum is complex-valued. We need to visualize the real and imaginary parts separately. By default, `plot_pairs()` plot the real part of the radially averaged spectral estimate:

```{r, fig.height = 9, fig.width = 9}
plot_pairs(est.list = KSDE, ppp = spp)
```

To examine the imaginary part, we set `type = "Im"`. As below figure shows, the values are almost zero, particularly for marginal cases where they should be exactly zero. Thus, the real part is our primary focus of analysis.

```{r, fig.height = 9, fig.width = 9}
plot_pairs(est.list = KSDE, ppp = spp, type = "Im")
```

# Coherence analysis

Similar to spectral analysis in time series, we can calculate the coherence and partial coherence based on the pseudo-spectrum for point processes. The `coherence()` function computes these measures for all non-overlaping frequencies and extracts their maxima. By default (`type = "partial"`), it returns the maximal partial coherence between any two point processes. In the code below, we calculate the maximal coherence and partial coherence for all pairwise point processes.

```{r}
coh.partial = coherence(sp.est = KSDE, ppp = spp) # Maximal partial coherence
coh = coherence(sp.est = KSDE, ppp = spp, type = "normal") # Maximal coherence
round(coh.partial, 2)
round(coh, 2)
```

Instead of summarizing by a single value, the `plot_coher()` function allows us to visualize the coherence and partial coherence values across all non-overlapping frequencies.

```{r, fig.height = 9, fig.width = 9}
plot_coher(sp.est = KSDE,
           coh.mat = coh,
           partial.coh.mat = coh.partial,
           ylim = c(0, 0.85))
```
